---
title: "Analysis of norming study (Sept 2018)"
author: '[Guillermo Montero-Melis](http://www.biling.su.se/montero_melis_guillermo)'
date: '`r as.character(format(Sys.Date(), format="%d/%m/%Y"))`'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    toc: yes
---

Introduction
============

Set up workspace
================

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library(GGally)
library(dplyr)
library(tidyr)
library(lme4)
knitr::opts_chunk$set(echo = FALSE, fig.height = 3.5)
```


```{r}
# load convenience functions
source("../Rfunctions/load_or_fit_fnc.R")
# some parameters
ggpairs_he <- 4.5  # default height for ggpairs plots
```


```{r, include = FALSE}
## load data files
# verb norming English
bias_en_l <- read.csv("data_verb-bias_L2eng.csv", fileEncoding = "UTF-8")
head(bias_en_l)
# verb norming Swedish
bias_sw_l <- read.csv("data_verb-bias_L1swe.csv", fileEncoding = "UTF-8")
head(bias_sw_l)
# verb comprehension/understanding
# compr <- read.csv(".csv",
#                   fileEncoding = "UTF-8")
# lextale task
# scored
lext <- read.csv("data_lextale_scored.csv", fileEncoding = "UTF-8")
head(lext)
# raw
lext_raw <- read.csv("data_lextale_raw.csv", fileEncoding = "UTF-8")
head(lext_raw)
# participant info
ppt <- read.csv("participant_info.csv", fileEncoding = "UTF-8")
head(ppt)
```


```{r}
# # In free translation task, consider translations scored as -99 (=unclear) as
# # incorrect (=0):
# und_free$score_unsimplified <- und_free$score  # keep copy of original variable
# und_free[und_free$score == -99, "score"] <- 0
```




Verb norming task: L2 English
==================================

The data from the verb norming task look like this:

```{r}
kable(head(bias_en_l[, 3:8]))
```

We have ratings for
`r length(unique(bias_en_l$verb))`
verbs, which we have collected on two different occasions. That means that
some verbs have been rated by more participants than others (see Appendix
for details), but this is not something we will worry about at first.


## Verb bias

For each verb, we can compute a single measure of *arm-minus-leg bias* (we call
it simply *bias*) by substracting the mean leg-relatedness rating from the mean
arm-relatedness rating for each verb, obtaining:

```{r}
bias_en <- bias_en_l %>%
  group_by(verb, category, rated_category) %>%
  summarise(rating = mean(rating),
            nObs = n()) %>%
  spread(rated_category, rating) %>%  # reshape to wide format for correlation plot
  mutate(bias = Arm - Leg)
kable(head(bias_en))
```

- A positive bias score (bias > 0) indicates bias towards arm-relatedness (max 6)
- A negative bias score (bias < 0) indicates bias towards leg-relatedness (min -6)
- A bias score = 0 indicates absence of either bias

```{r}
# order factor levels of verbs depending on their bias (for plotting)
ordered_verbs_bias <- as.character(pull(bias_en[order(bias_en$bias), ], verb))
bias_en$verb <- factor(bias_en$verb, levels = ordered_verbs_bias)
```


```{r, fig.height = 12}
ggplot(bias_en, aes(x = verb, y = bias, colour = category)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  theme(axis.text.x = element_text(size = 5)) +
  coord_flip() +
  ggtitle("Mean arm/leg bias of verbs")
```

The figure shows that the bias was generally stronger for the verbs we had
categorized as arm verbs than for those we thought of as leg verbs.

This can be shown even clearer if we just consider the size of the bias in
the expected direction, so a negative bias now constitutes a bias in the 
"wrong" direction (e.g., an arm bias for a supposedly leg-related verb).
The following plot shows this:

```{r}
# We add a bias column that changes the sign for leg verbs, so that bias
# is expressed with respect to the expected direction of the bias for a verb
bias_en$bias_absol <- bias_en$bias
bias_en[bias_en$category == "leg", "bias_absol"] <- - bias_en[bias_en$category == "leg", "bias"]
ggplot(bias_en, aes(x = category, y = bias_absol, colour = category)) + 
  geom_boxplot() +
  ylab("Size of bias in predicted direction") +
  ggtitle("Size of congruent bias of English verbs")
```

Among the English verbs, `r sum(bias_en$bias_absol < 0)` had opposite bias
to what was expected:

```{r}
kable(bias_en[bias_en$bias_absol < 0, ])
```

To scuff means to "Scrape or brush the surface of (a shoe or other object) 
against something" (OED).
But participants seemed not to know the verb very well and assume it had some
arm-related meaning.

To dash means (OED):

1. Run or travel somewhere in a great hurry.
2. Strike or fling (something) somewhere with great force, especially so as to
have a destructive effect; hurl.

In the second sense it is considered a synonym of, among others,
*hurl, smash, crash, slam, throw, toss*, all of which are hand-related, so it
is in fact understandable that it had a arm/hand bias

```{r}
# Save the dataframe containing verb biases (`bias_en`) to disk:
write.csv(bias_en, "verb-bias_en_mean.csv", row.names = FALSE, fileEncoding = "UTF-8")
```



## Cut-off at bias = 3

Somewhat arbitrarily, we could say that a bias of 3 in the expected direction
constitutes an acceptable cut-off point to select a verb for the study.
There are 
`r sum(with(bias_en, bias_absol >= 3 & category == "arm"))`
arm verbs (out of `r sum(with(bias_en, category == "arm"))`) that satisfy this condition,
but only
`r sum(with(bias_en, bias_absol >= 3 & category == "leg"))`
leg verbs (out of
`r sum(with(bias_en, category == "leg"))`) that qualify.

Here are the leg verbs that have a clear bias:

```{r, fig.height = 5}
ggplot(bias_en[with(bias_en, bias_absol >= 3 & category == "leg"), ],
       aes(x = verb, y = bias)) +
  geom_point() +
  theme(axis.text.x = element_text(size = 6)) +
  coord_flip() +
  ggtitle("English leg verbs with strongest leg bias")
```


## Spread of ratings

Above we have only considered the mean rating for each verb (averaged across
participants).
But how much spread was there in the ratings from individual participants?

```{r}
bias_spread_en <- bias_en_l %>%
  select(participant, verb:rating) %>%
  spread(rated_category, rating) %>%
  mutate(bias = Arm - Leg)
# order verbs according to bias
bias_spread_en$verb <- factor(bias_spread_en$verb, levels = ordered_verbs_bias)
```


```{r, fig.height = 12}
ggplot(bias_spread_en, aes(x = verb, y = bias, colour = category)) +
  geom_hline(yintercept = 0) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(alpha = .4) +
  ylab("Verb bias") + 
  coord_flip() +
  ggtitle("Variability around mean arm/leg bias of English verbs")
```

What this figure shows is that the average bias for a verb can still hide some
important individual differences.
For instance, the leg verb *kneel* had a fairly strong leg bias of
`r round(bias_en[bias_en$verb == "kneel", ]$bias, 1)`
and still there was one participant who rated it as completely arm-biased
(bias = 6).

Some of these data points might simply be due to participants getting the 
scales mixed up, but probably not all of these data points came about in this
way:
Consider the supposedly leg verbs *tread* and *bounce*.
For the latter, one could imagine how a participant made associations with 
*bouncing a ball*, in which case the verb *to bounce* naturally becomes 
arm-related).

Finally, note also the many ratings of verbs as having no bias (verb bias = 0).
These may be the result of participants not knowing a verb meaning,
a matter to which we now turn.



Verb norming task: L1 Swedish
=============================

The data from the Swedish verb norming task look like this:

```{r}
kable(head(bias_sw_l[, 3:8]))
```

We have ratings for
`r length(unique(bias_sw_l$verb))`
verbs, all of which were collected in Sept 2018. We had
`r length(unique(bias_sw_l$participant))`
native Swedish speakers who provided the ratings.



## Verb bias

For each verb, we can compute a single measure of *arm-minus-leg bias* (we call
it simply *bias*) by substracting the mean leg-relatedness rating from the mean
arm-relatedness rating for each verb, obtaining:

```{r}
bias_sw <- bias_sw_l %>%
  group_by(verb, category, rated_category) %>%
  summarise(rating = mean(rating)) %>%
  spread(rated_category, rating) %>%  # reshape to wide format for correlation plot
  mutate(bias = Arm - Leg)
kable(head(bias_sw))
```

- A positive bias score (bias > 0) indicates bias towards arm-relatedness (max 6)
- A negative bias score (bias < 0) indicates bias towards leg-relatedness (min -6)
- A bias score = 0 indicates absence of either bias

```{r}
# order factor levels of verbs depending on their bias (for plotting)
ordered_verbs_bias <- as.character(pull(bias_sw[order(bias_sw$bias), ], verb))
bias_sw$verb <- factor(bias_sw$verb, levels = ordered_verbs_bias)
```


```{r, fig.height = 12}
ggplot(bias_sw, aes(x = verb, y = bias, colour = category)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  theme(axis.text.x = element_text(size = 5)) +
  coord_flip() +
  ggtitle("Mean arm/leg bias of Swedish verbs")
```

The figure shows that the bias was generally stronger for the verbs we had
categorized as arm verbs than for those we thought of as leg verbs.

This can be shown even clearer if we just consider the size of the bias in
the expected direction, so a negative bias now constitutes a bias in the 
"wrong" direction (e.g., an arm bias for a supposedly leg-related verb).
The following plot shows this:

```{r}
# We add a bias column that changes the sign for leg verbs, so that bias
# is expressed with respect to the expected direction of the bias for a verb
bias_sw$bias_absol <- bias_sw$bias
bias_sw[bias_sw$category == "leg", "bias_absol"] <- - bias_sw[bias_sw$category == "leg", "bias"]
ggplot(bias_sw, aes(x = category, y = bias_absol, colour = category)) + 
  geom_boxplot() +
  ylab("Size of bias in predicted direction") +
  ggtitle("Size of congruent bias of Swedish verbs")
```

Among the Swedish verbs, `r sum(bias_sw$bias_absol < 0)` had opposite
bias to what was expected:

```{r}
kable(bias_sw[bias_sw$bias_absol < 0, ])
```

All of these verbs are leg-related.

```{r}
# Save the dataframe containing verb biases (`bias_sw`) to disk:
write.csv(bias_sw, "verb-bias_sw_mean.csv", row.names = FALSE, fileEncoding = "UTF-8")
```



## Cut-off at bias = 3

Somewhat arbitrarily, we could say that a bias of 3 in the expected direction
constitutes an acceptable cut-off point to select a verb for the study.
There are 
`r sum(with(bias_sw, bias_absol >= 3 & category == "arm"))`
arm verbs (out of `r sum(with(bias_sw, category == "arm"))`) that satisfy this condition,
but only
`r sum(with(bias_sw, bias_absol >= 3 & category == "leg"))`
leg verbs (out of
`r sum(with(bias_sw, category == "leg"))`) that qualify.

Here are the leg verbs that have a clear bias:

```{r, fig.height = 5}
ggplot(bias_sw[with(bias_sw, bias_absol >= 3 & category == "leg"), ],
       aes(x = verb, y = bias)) +
  geom_point() +
  theme(axis.text.x = element_text(size = 6)) +
  coord_flip() +
  ggtitle("Swedish leg verbs with strongest leg bias")
```


## Spread of ratings

Above we have only considered the mean rating for each verb (averaged across
participants).
But how much spread was there in the ratings from individual participants?

```{r}
bias_spread_sw <- bias_sw_l %>%
  select(participant, verb:rating) %>%
  spread(rated_category, rating) %>%
  mutate(bias = Arm - Leg)
# order verbs according to bias
bias_spread_sw$verb <- factor(bias_spread_sw$verb, levels = ordered_verbs_bias)
```


```{r, fig.height = 12}
ggplot(bias_spread_sw, aes(x = verb, y = bias, colour = category)) +
  geom_hline(yintercept = 0) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(alpha = .4) +
  ylab("Verb bias") + 
  coord_flip() +
  ggtitle("Variability around mean arm/leg bias of Swedish verbs")
```

What this figure shows is that the average bias for a verb can still hide some
important individual differences.
Some of these data points might simply be due to participants getting the 
scales mixed up, but probably not all of these data points came about in this
way.

Note also the some ratings have no bias (verb bias = 0). This is unlikely to
be due to Swedish speakers not knowing the verbs (since they were native 
speakers).


L2 verb comprehension task
======================



Choice of target verbs
======================

English verbs
-------------

As a first approximation, let's just choose the 40 verbs in each category with
the strongest (absolute) bias.






Swedish verbs
-------------



Appendices
==========

Full list of normed L2 verbs (June and September)
------------------------------------------------

List of all normed L2 English verbs, sorted after bias:

```{r}
kable(bias_en %>% arrange(bias, verb))
```

The column nObs (=number of observations) shows how many participants rated
a given verb.

Note that we had verbs rated for their arm- vs leg-relatedness on two separate
occasions: in the pilot study in June, and in the norming study in September
2018. Most of the verbs were rated on both occasions, but some were just
rated either in June or in September.

```{r}
vbs_junsep <- sort(unique(bias_en_l$verb[bias_en_l$verb_normed == "june+sept"]))
vbs_jun <- sort(unique(bias_en_l$verb[bias_en_l$verb_normed == "june"]))
vbs_sep <- sort(unique(bias_en_l$verb[bias_en_l$verb_normed == "sept"]))
```


Verbs rated **both in June and in September**:

`r vbs_junsep`


Verbs rated **in June only**:

`r vbs_jun`


Verbs rated **in September only**:

`r vbs_sep`



Main results and take-away messages
========================

## Conclusions from L2 English verb norming

Participants assessed the strength of arm and leg association for each verb.
We computed the bias by substracting these two ratings ($arm - leg$).



## Conclusions from L1 Swedish verb norming


## Conclusions from verb comprehension task

Participants had to translate L2 English verbs to their L1 Swedish.
Each verb was basically rated as an either correct or incorrect translation.
However, we did also consider intermediate cases of underspecified translations
(e.g., *gå* for 'stroll').

Conclusions:



## Overall suggestions based on norming study


