---
title: "Analysis of norming study (Sept 2018)"
author: '[Guillermo Montero-Melis](http://www.biling.su.se/montero_melis_guillermo)'
date: '`r as.character(format(Sys.Date(), format="%d/%m/%Y"))`'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    toc: yes
---

Introduction
============

Set up workspace
================

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library(GGally)
library(dplyr)
library(tidyr)
library(lme4)
knitr::opts_chunk$set(echo = TRUE, fig.height = 3.5)
```


```{r}
# load convenience functions
source("../Rfunctions/load_or_fit_fnc.R")
# some parameters
ggpairs_he <- 4.5  # default height for ggpairs plots
```


```{r, include = TRUE}
## load data files
# verb norming English
bias_en_l <- read.csv("data_verb-bias_L2eng.csv", fileEncoding = "UTF-8")
head(bias_en_l)
# verb norming Swedish
bias_sw_l <- read.csv("data_verb-bias_L1swe.csv", fileEncoding = "UTF-8")
head(bias_sw_l)
# verb comprehension/understanding
# compr <- read.csv(".csv",
#                   fileEncoding = "UTF-8")
# lextale task
# scored
lext <- read.csv("data_lextale_scored.csv", fileEncoding = "UTF-8")
head(lext)
# raw
lext_raw <- read.csv("data_lextale_raw.csv", fileEncoding = "UTF-8")
head(lext_raw)
# participant info
ppt <- read.csv("participant_info.csv", fileEncoding = "UTF-8")
head(ppt)
```


```{r}
# # In free translation task, consider translations scored as -99 (=unclear) as
# # incorrect (=0):
# und_free$score_unsimplified <- und_free$score  # keep copy of original variable
# und_free[und_free$score == -99, "score"] <- 0
```




Verb norming task: L2 English
==================================

The data from the verb norming task look like this:

```{r}
kable(head(bias_en_l[, 3:8]))
```

We have ratings for
`r length(unique(bias_en_l$verb))`
verbs, which we have collected on two different occasions. That means that
some verbs have been rated by more participants than others (see Appendix
for details), but this is not something we will worry about at first.


## Verb bias

For each verb, we can compute a single measure of *arm-minus-leg bias* (we call
it simply *bias*) by substracting the mean leg-relatedness rating from the mean
arm-relatedness rating for each verb, obtaining:

```{r}
bias_en <- bias_en_l %>%
  group_by(verb, category, rated_category) %>%
  summarise(rating = mean(rating),
            nObs = n()) %>%
  spread(rated_category, rating) %>%  # reshape to wide format for correlation plot
  mutate(bias = Arm - Leg)
kable(head(bias_en))
```

- A positive bias score (bias > 0) indicates bias towards arm-relatedness (max 6)
- A negative bias score (bias < 0) indicates bias towards leg-relatedness (min -6)
- A bias score = 0 indicates absence of either bias

```{r}
# order factor levels of verbs depending on their bias (for plotting)
ordered_verbs_bias <- as.character(pull(bias_en[order(bias_en$bias), ], verb))
bias_en$verb <- factor(bias_en$verb, levels = ordered_verbs_bias)
```


```{r, fig.height = 12}
ggplot(bias_en, aes(x = verb, y = bias, colour = category)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  theme(axis.text.x = element_text(size = 5)) +
  coord_flip() +
  ggtitle("Mean arm/leg bias of verbs")
```

The figure shows that the bias was generally stronger for the verbs we had
categorized as arm verbs than for those we thought of as leg verbs.

This can be shown even clearer if we just consider the size of the bias in
the expected direction, so now a negative bias constitutes a bias in the
"wrong" direction (e.g., an arm bias for a supposedly leg-related verb).
The following plot shows this:

```{r}
# We add a bias column that changes the sign for leg verbs, so that bias
# is expressed with respect to the expected direction of the bias for a verb
bias_en$bias_absol <- bias_en$bias
bias_en[bias_en$category == "leg", "bias_absol"] <- - bias_en[bias_en$category == "leg", "bias"]
ggplot(bias_en, aes(x = category, y = bias_absol, colour = category)) + 
  geom_boxplot() +
  ylab("Size of bias in predicted direction") +
  ggtitle("Size of congruent bias of English verbs")
```

Among the English verbs, `r sum(bias_en$bias_absol < 0)` had opposite bias
to what was expected:

```{r}
kable(bias_en[bias_en$bias_absol < 0, ])
```

To scuff means to "Scrape or brush the surface of (a shoe or other object) 
against something" (OED).
But participants seemed not to know the verb very well and assume it had some
arm-related meaning.

To dash means (OED):

1. Run or travel somewhere in a great hurry.
2. Strike or fling (something) somewhere with great force, especially so as to
have a destructive effect; hurl.

In the second sense it is considered a synonym of, among others,
*hurl, smash, crash, slam, throw, toss*, all of which are hand-related, so it
is in fact understandable that it had a arm/hand bias

```{r}
# Save the dataframe containing verb biases (`bias_en`) to disk:
write.csv(bias_en, "verb-bias_en_mean.csv", row.names = FALSE, fileEncoding = "UTF-8")
```



## Cut-off at bias = 3

Somewhat arbitrarily, we could say that a bias of 3 in the expected direction
constitutes an acceptable cut-off point to select a verb for the study.
There are 
`r sum(with(bias_en, bias_absol >= 3 & category == "arm"))`
arm verbs (out of `r sum(with(bias_en, category == "arm"))`) that satisfy this condition,
but only
`r sum(with(bias_en, bias_absol >= 3 & category == "leg"))`
leg verbs (out of
`r sum(with(bias_en, category == "leg"))`) that qualify.

Here are the leg verbs that have a clear bias:

```{r, fig.height = 5}
ggplot(bias_en[with(bias_en, bias_absol >= 3 & category == "leg"), ],
       aes(x = verb, y = bias)) +
  geom_point() +
  theme(axis.text.x = element_text(size = 6)) +
  coord_flip() +
  ggtitle("English leg verbs with strongest leg bias")
```


## Spread of ratings

Above we have only considered the mean rating for each verb (averaged across
participants).
But how much spread was there in the ratings from individual participants?

```{r}
bias_spread_en <- bias_en_l %>%
  select(participant, verb:rating) %>%
  spread(rated_category, rating) %>%
  mutate(bias = Arm - Leg)
# order verbs according to bias
bias_spread_en$verb <- factor(bias_spread_en$verb, levels = ordered_verbs_bias)
```


```{r, fig.height = 12}
ggplot(bias_spread_en, aes(x = verb, y = bias, colour = category)) +
  geom_hline(yintercept = 0) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(alpha = .4) +
  ylab("Verb bias") + 
  coord_flip() +
  ggtitle("Variability around mean arm/leg bias of English verbs")
```

What this figure shows is that the average bias for a verb can still hide some
important individual differences.
For instance, the leg verb *kneel* had a fairly strong leg bias of
`r round(bias_en[bias_en$verb == "kneel", ]$bias, 1)`
and still there was one participant who rated it as completely arm-biased
(bias = 6).

Some of these data points might simply be due to participants getting the 
scales mixed up, but probably not all of these data points came about in this
way:
Consider the supposedly leg verbs *tread* and *bounce*.
For the latter, one could imagine how a participant made associations with 
*bouncing a ball*, in which case the verb *to bounce* naturally becomes 
arm-related).

Finally, note also the many ratings of verbs as having no bias (verb bias = 0).
These may be the result of participants not knowing a verb meaning,
a matter to which we now turn.



Verb norming task: L1 Swedish
=============================

The data from the Swedish verb norming task look like this:

```{r}
kable(head(bias_sw_l[, 3:8]))
```

We have ratings for
`r length(unique(bias_sw_l$verb))`
verbs, all of which were collected in Sept 2018. We had
`r length(unique(bias_sw_l$participant))`
native Swedish speakers who provided the ratings.



## Verb bias

For each verb, we can compute a single measure of *arm-minus-leg bias* (we call
it simply *bias*) by substracting the mean leg-relatedness rating from the mean
arm-relatedness rating for each verb, obtaining:

```{r}
bias_sw <- bias_sw_l %>%
  group_by(verb, category, rated_category) %>%
  summarise(rating = mean(rating)) %>%
  spread(rated_category, rating) %>%  # reshape to wide format for correlation plot
  mutate(bias = Arm - Leg)
kable(head(bias_sw))
```

- A positive bias score (bias > 0) indicates bias towards arm-relatedness (max 6)
- A negative bias score (bias < 0) indicates bias towards leg-relatedness (min -6)
- A bias score = 0 indicates absence of either bias

```{r}
# order factor levels of verbs depending on their bias (for plotting)
ordered_verbs_bias <- as.character(pull(bias_sw[order(bias_sw$bias), ], verb))
bias_sw$verb <- factor(bias_sw$verb, levels = ordered_verbs_bias)
```


```{r, fig.height = 12}
ggplot(bias_sw, aes(x = verb, y = bias, colour = category)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  theme(axis.text.x = element_text(size = 5)) +
  coord_flip() +
  ggtitle("Mean arm/leg bias of Swedish verbs")
```

The figure shows that the bias was generally stronger for the verbs we had
categorized as arm verbs than for those we thought of as leg verbs.

This can be shown even clearer if we just consider the size of the bias in
the expected direction, so a negative bias now constitutes a bias in the 
"wrong" direction (e.g., an arm bias for a supposedly leg-related verb).
The following plot shows this:

```{r}
# We add a bias column that changes the sign for leg verbs, so that bias
# is expressed with respect to the expected direction of the bias for a verb
bias_sw$bias_absol <- bias_sw$bias
bias_sw[bias_sw$category == "leg", "bias_absol"] <- - bias_sw[bias_sw$category == "leg", "bias"]
ggplot(bias_sw, aes(x = category, y = bias_absol, colour = category)) + 
  geom_boxplot() +
  ylab("Size of bias in predicted direction") +
  ggtitle("Size of congruent bias of Swedish verbs")
```

Among the Swedish verbs, `r sum(bias_sw$bias_absol < 0)` had opposite
bias to what was expected:

```{r}
kable(bias_sw[bias_sw$bias_absol < 0, ])
```

All of these verbs are leg-related.

```{r}
# Save the dataframe containing verb biases (`bias_sw`) to disk:
write.csv(bias_sw, "verb-bias_sw_mean.csv", row.names = FALSE, fileEncoding = "UTF-8")
```



## Cut-off at bias = 3

Somewhat arbitrarily, we could say that a bias of 3 in the expected direction
constitutes an acceptable cut-off point to select a verb for the study.
There are 
`r sum(with(bias_sw, bias_absol >= 3 & category == "arm"))`
arm verbs (out of `r sum(with(bias_sw, category == "arm"))`) that satisfy this condition,
but only
`r sum(with(bias_sw, bias_absol >= 3 & category == "leg"))`
leg verbs (out of
`r sum(with(bias_sw, category == "leg"))`) that qualify.

Here are the leg verbs that have a clear bias:

```{r, fig.height = 5}
ggplot(bias_sw[with(bias_sw, bias_absol >= 3 & category == "leg"), ],
       aes(x = verb, y = bias)) +
  geom_point() +
  theme(axis.text.x = element_text(size = 6)) +
  coord_flip() +
  ggtitle("Swedish leg verbs with strongest leg bias")
```


## Spread of ratings

Above we have only considered the mean rating for each verb (averaged across
participants).
But how much spread was there in the ratings from individual participants?

```{r}
bias_spread_sw <- bias_sw_l %>%
  select(participant, verb:rating) %>%
  spread(rated_category, rating) %>%
  mutate(bias = Arm - Leg)
# order verbs according to bias
bias_spread_sw$verb <- factor(bias_spread_sw$verb, levels = ordered_verbs_bias)
```


```{r, fig.height = 12}
ggplot(bias_spread_sw, aes(x = verb, y = bias, colour = category)) +
  geom_hline(yintercept = 0) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(alpha = .4) +
  ylab("Verb bias") + 
  coord_flip() +
  ggtitle("Variability around mean arm/leg bias of Swedish verbs")
```

What this figure shows is that the average bias for a verb can still hide some
important individual differences.
Some of these data points might simply be due to participants getting the 
scales mixed up, but probably not all of these data points came about in this
way.

Note also the some ratings have no bias (verb bias = 0). This is unlikely to
be due to Swedish speakers not knowing the verbs (since they were native 
speakers).


L2 verb comprehension task
======================

In the norming study in Sept 2018, 12 Swedish participants carried out a
comprehension task for the L2 English verbs: they had to translate the English
verbs to Swedish. The translations were then scored as either correct or
incorrect by two Swedish native speakers (see 
`norming_1809_analysis/scoring_translation-task.Rmd` for details about inter-
rater agreement).


The data set and some processing
--------------------------------

The final data set (after resolving disagreements between raters) looks like
this:

```{r}
compr <- read.csv("data_translations_scored.csv", fileEncoding = "UTF-8",
                  stringsAsFactors = FALSE)
kable(head(compr %>% select(participant:cognate)))
```

Note that some verbs haven't any verb_category assigned:

```{r}
# Note that some verbs haven't any verb_category assigned:
unique(compr %>% select(expName, verb, verb_category:in_orig)) %>% 
  filter(is.na(verb_category)) %>% kable
```

These are verbs for which we collected data in June, but which we subsequently
excluded because they were unsuitable for several reasons (ambiguous, unknown,
low bias...). So simply remove them from the data set. (We will remove
`r sum(is.na(compr$verb_category))` observations.)

```{r}
compr <- compr[! is.na(compr$verb_category), ]
```



The coding scheme was:

- 1 = Correct translation
- 0 = Incorrect translation
- 0.5 = Partly correct but underspecified (e.g., 'gå' for 'lumber')
- -99 = Undecided

Here's a table over how much each of the scoring occurred:

```{r}
table(compr$score)
```

We will remove scores of -99 (treat them as NAs) and leave the rest as they are.

```{r}
compr <- compr[! compr$score == -99, ]
```


Create a data frame with the mean score per verb (expressed as percentage of
number of responses provided)

```{r}
# Create a data frame with the mean ratings
(compr_vbs <- compr %>%
  group_by(verb, verb_category, cognate_bin, cognate) %>%
  summarise(sumScore = sum(score),
            N = n(),
            Score = sumScore / N)) %>%
  head %>% kable
```


Verb comprehension by word type
-------------------------------

Show Scores by verb type:

```{r}
ggplot(compr_vbs, aes(x = verb_category, y = Score, colour = verb_category)) +
  geom_boxplot() + ylab("Proportion correct translations")
```

Leg verbs were less well understood than arm verbs.

Let's look at binned scores:

```{r}
compr_vbs$binnedScore <- cut(compr_vbs$Score, seq(0, 1, .25), right = FALSE, 
                             include.lowest = TRUE)
with(compr_vbs, table(binnedScore, verb_category))
```

We should probably only include verbs that have at least a mean score of 50%
correct translations.


Choice of target verbs
======================

We need to take into account:

- Verb bias (L2 Eng and Swe)
- Verb comprehension (L2 Eng)
- Whether it's a translation equivalent (Eng-to-Swe)
- Cognate status (L2 Engl)
- Other frequency and word length measures


English verbs
-------------

The two most important factors are verb bias and whether it's understood
by Swedish speakers. Let's combine that information:

```{r}
# Add info about verb bias to the comprehension scores
compr_vbs <- left_join(compr_vbs, bias_en %>% select(verb, bias, bias_absol))
```

And plot it:

```{r}
ggplot(compr_vbs, aes(x = Score, y = bias_absol, colour = verb_category)) +
  geom_point() +
  geom_vline(xintercept = .5, linetype = "dashed") +
  geom_hline(yintercept = 3, linetype = "dashed") +
  ylab("Verb bias in coherent direction") +
  xlab("Proportion correct translations") +
  ggtitle("Bias and translation accuracy for L2 English verbs")
```

Ideally we want to keep only verbs that are in the upper right quadrant. These
are verbs that both were translated correctly at least 50% of the time and
elicited a bias of at least 3 in the expected direction.


### First selection of L2 English verbs

Let's take a look at them:

```{r}
# Subset verbs that meet criteria, but relax bias to catch a verb that is right
# below 3 (it's trek, with bias 2.92, and switch, with bias 2.97)
select_en <- compr_vbs %>% 
  filter(Score >= .5 & bias_absol >= 2.9)
```

```{r, fig.height = 6, fig.width = 10}
ggplot(select_en, aes(x = Score, y = bias_absol, colour = verb_category,
                      label = verb)) +
  geom_text() +
  ylab("Verb bias in coherent direction") +
  xlab("Proportion correct translations") +
  ggtitle("First selection of L2 English verbs")
```

How many of each type are there?

```{r}
table(select_en$verb_category)
```


### Can we squeeze out a few more leg verbs

We are short of 3 leg verbs to have 10 quadruples per block (which would
require 40 distinct verbs).

To see if we can find a few more leg verbs We can relax our requirements
somewhat for leg verbs. So let's take a look at verbs in the upper left and
lower right quadrants from the figure above.

Verbs with high leg bias but poor translation accuracy, or the opposite, verbs
with high translation accuracyl but low leg bias:

```{r, fig.height = 5}
compr_vbs %>% 
  filter(verb_category == "leg" & (Score > .5 & bias_absol < 3)
         | (Score < .5 & bias_absol > 3)) %>%
  ggplot(aes(x = Score, y = bias_absol, label = verb)) +
  geom_text() +
  geom_vline(xintercept = .5, linetype = "dashed") +
  geom_hline(yintercept = 3, linetype = "dashed") +
  xlim(0,1) + ylim(0,6) +
  ylab("Verb bias in coherent direction") +
  xlab("Proportion correct translations") 
```

We can include the three verbs that were well understood and elicited a
reasonable bias: creep, dance, and stamp.


```{r}
select_en <- rbind(select_en, 
                   compr_vbs[compr_vbs$verb %in% c("creep", "dance", "stamp"),])
kable(tail(select_en))
```


### Check other variables

Check cognate status

```{r}
with(select_en, addmargins(table(cognate_bin, verb_category)))
```


Number of letters:

```{r}
select_en$nbLetters <- nchar(select_en$verb)
with(select_en, addmargins(table(nbLetters, verb_category)))
select_en %>% group_by(verb_category) %>%
  summarise(M_letters = mean(nbLetters),
            SD_letters = sd(nbLetters),
            min_letters = min(nbLetters),
            max_letters = max(nbLetters))
```




### Temporary solution


```{r}
# Number of verbs to choose per category
N_verbs_per_categ <- 40
```

As a first approximation, let's just choose the `r N_verbs_per_categ`
verbs in each category with the strongest (absolute) bias.

```{r}
N_verbs_per_categ <- 40
en_arm <- bias_en[bias_en$category == "arm", ] %>%
  arrange(-bias_absol, nObs, verb) %>%
  select(category, verb)
en_leg <- bias_en[bias_en$category == "leg", ] %>%
  arrange(-bias_absol, nObs, verb) %>%
  select(category, verb)
# Choose the ones with the strongest bias in each category
en_targets <- rbind(en_arm[1:N_verbs_per_categ,], en_leg[1:N_verbs_per_categ,])
en_targets <- en_targets %>%
  ungroup() %>%
  mutate(verb = as.character(verb)) %>%  # to order verbs alphabetically
  arrange(category, verb) %>%
  rename(type = category)  # for compatibility with functions in other scripts
head(en_targets); tail(en_targets)
# Save to disk (two different destinations):
write.csv(en_targets, "english-targets.csv", row.names = FALSE,
          fileEncoding = "UTF-8")
write.csv(en_targets, "../exp-scripts_psychopy/replication_pilot_1810/stimuli/english-targets.csv",
          row.names = FALSE, fileEncoding = "UTF-8")
```



Swedish verbs
-------------



Appendices
==========

Full list of normed L2 verbs (June and September)
------------------------------------------------

List of all normed L2 English verbs, sorted after bias, and adding translation
scores if available:

```{r}
bias_en %>% 
  left_join(compr_vbs %>% select(verb, Score)) %>%
  arrange(bias, verb) %>%
  kable
```

The column nObs (=number of observations) shows how many participants rated
a given verb.

Note that we had verbs rated for their arm- vs leg-relatedness on two separate
occasions: in the pilot study in June, and in the norming study in September
2018. Most of the verbs were rated on both occasions, but some were just
rated either in June or in September.

```{r}
vbs_junsep <- sort(unique(bias_en_l$verb[bias_en_l$verb_normed == "june+sept"]))
vbs_jun <- sort(unique(bias_en_l$verb[bias_en_l$verb_normed == "june"]))
vbs_sep <- sort(unique(bias_en_l$verb[bias_en_l$verb_normed == "sept"]))
```


Verbs rated **both in June and in September**:

`r vbs_junsep`


Verbs rated **in June only**:

`r vbs_jun`


Verbs rated **in September only**:

`r vbs_sep`



Main results and take-away messages
========================

## Conclusions from L2 English verb norming

Participants assessed the strength of arm and leg association for each verb.
We computed the bias by substracting these two ratings ($arm - leg$).



## Conclusions from L1 Swedish verb norming


## Conclusions from verb comprehension task

Participants had to translate L2 English verbs to their L1 Swedish.
Each verb was basically rated as an either correct or incorrect translation.
However, we did also consider intermediate cases of underspecified translations
(e.g., *gå* for 'stroll').

Conclusions:



## Overall suggestions based on norming study


