---
title: "Execution of paradiddles"
author: '[Guillermo Montero-Melis](https://www.mpi.nl/people/montero-melis-guillermo)'
date: '`r as.character(format(Sys.Date(), format="%d/%m/%Y"))`'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Introduction
============

Check visually how the paradiddles are executed. This is useful to test the
drumpads and check they are registering the taps.


Setup workspace
===============

Libraries:

```{r, message=FALSE, warning=FALSE}
library("knitr")
library("tidyverse")  # ggplot2, dplyr, readr, purrr, etc
theme_set(theme_bw())
```


Import data (pick log file to be inspected from menu):

```{r}
# data file
f <- file.choose()
# f <- "data/998_sheb_replication_1810_2020_Mar_13_1340.log"
ncols <- max(count.fields(f, sep = "\t"))
logf <- read.delim(
  f, sep = "\t", header = FALSE,
  col.names = c("t", "type", paste("msg", seq_len(ncols-2), sep = ""))
  )
print(f)
```

```{r}
kable(head(logf))
```

Extract info
=============

Now let's extract the info we need from the log file.

Order of blocks:

```{r}
# Order in which conditions were carried out
get_order <- function(logf) {
  order_line <- logf$msg1[grep("Imported block_order_", logf$msg1)]
  order <- gsub("Imported block_order_(.*)\\.csv.*", "\\1", order_line)
  tibble(block = 1 : 3, condition = unlist(str_split(order, "-")))
}
(cond_order <- get_order(logf)) %>% kable
```


Block and trial structure:

```{r}
# Get the overall block and trial structure
get_structure <- function(df) {
  
  # Description of the upcoming trial
  trials <- df[grepl("start of memory period", df$msg1), ]
  # Lines containing timing for the start of memory period
  start <- df[grepl("b_memory_period_2_: autoDraw = True", df$msg1), ]
  # Lines containing timing for the *end* of memory period:
  # Function looks for end of the period, but keeps only 1st of 2 repeated lines
  mem_end <- function (df) {
    end_Ls <- grep("b_memory_period_2_: autoDraw = False", df$msg1)
    keep <- rep(c(TRUE, FALSE), length.out = length(end_Ls))
    end_Ls[keep]
  }
  end <- df[mem_end(df), ]
  
  # Sanity check there should be one start/end timing for each trial:
  if (! ( nrow(trials) == nrow(start) & nrow(trials) == nrow(end) )) {
    stop("Number of trials and timings don't match!")
  } 

  tibble(
    routine = gsub("(.*)/([0-9])/([0-9]+)/.*", "\\1", trials$msg1),
    block   = as.numeric(gsub("(.*)/([0-9])/([0-9]+)/.*", "\\2", trials$msg1)),
    trial   = as.numeric(gsub("(.*)/([0-9])/([0-9]+)/.*", "\\3", trials$msg1)),
    t0      = start$t,
    tend    = end$t,
    # 1st and last lines that contain paradiddle info if carried out:
    L0      = as.numeric(row.names(start)) + 1,
    Lend    = as.numeric(row.names(end)) - 1
  )
}
(mystr <- get_structure(logf)) %>% head %>% kable
mystr %>% tail %>% kable
```


Function to extract the relevant tapping info:

```{r}
# as input it takes a dataframe, information about the block and trial
# it is in, the first and last Lines where it should look for taps, etc.
get_taps <- function(df, routine, block, trial,  t0, L0, Lend, ...) {
  # Look only in rows specified as argument
  df <- df[L0 : Lend, ]
  # Select only relevant rows but remove rows for when the tap stopped (velocity=0)
  taps <- df[grepl("SamplePad", df$msg1) & !grepl("velocity=0", df$msg1), ]
  # Extract the note played at a point in time and include block+trial info
  out <- if (nrow(taps) == 0) {  # if no taps in the trial
    tibble(
      routine,
      block,
      trial,
      time_exp = NA,
      time     = NA,
      note     = NA,
      pad      = NA
      )
    } else {
      tibble(
        routine = routine,
        block,
        trial,
        time_exp = taps$t,
        time     = time_exp - t0,  # time within memory phase
        note     = as.numeric(gsub(".*note=([0-9]+).*", "\\1", taps$msg1)),
        pad      = gsub(".*/(SamplePad [0-9])/.*", "\\1", taps$msg1)
      )
    }
  out
}
```


Put it all together in a data frame:

```{r}
taps <- pmap(mystr, get_taps, df = logf) %>% bind_rows()
taps %>% head() %>% kable()
```


Do the mapping between notes and effectors
==================================

```{r}
# This bit needs to be tested more thoroughly. It assumes that the notes and
# SamplePad IDs are constant and assigned consistently across participants.

(note_mappings <- tibble(
  note     = c(45, 48, 49, 51),
  pad      = rep(paste("SamplePad", c(0, 1)), each = 2),
  side     = rep(c("R", "L"), 2),
  effector = rep(c("hands", "feet"), each = 2)
)) %>% kable()

# Join info:

taps <- taps %>%
  left_join(cond_order) %>%
  left_join(note_mappings)

taps %>% head() %>% kable()
```



Visualize
=========

```{r}
# Function to plot different routines:
plot_taps <- function (df, rout, nomapping = TRUE)  {
  df <- df %>% 
    filter(routine == rout) %>%
    mutate(note = factor(note))
  if (rout == "word_presentation_practice") {
    p <- ggplot(df) +
      geom_vline(aes(xintercept = time, linetype = effector), size = 1.5)
  } else {
    if (nomapping) {
      p <- ggplot(df) +
        geom_vline(aes(xintercept = time, colour = note), size = 1)
    } else if (! nomapping) {
        p <- ggplot(df) +
          geom_vline(
              aes(xintercept = time, colour = side, linetype = effector),
              size = 1)
    } else {
          stop("nomapping must be logical")
    }
    p <- p + facet_grid(trial ~ condition)
  }
  p + geom_vline(xintercept = 3) + ggtitle(rout)
}
```



For the first practice trials there should be no tapping, so if this plot is not
empty there is a problem:

```{r, fig.height=2}
plot_taps(taps, "word_presentation_practice")
```


Block-specific training trials:

```{r}
plot_taps(taps, "word_presentation_training")
```

Real trials:

```{r, fig.height=10}
plot_taps(taps, "word_presentation")
```


```{r, fig.height=10}
# Visualize with mapping
# plot_taps(taps, "word_presentation", nomapping = FALSE)
```
