---
title: "Analysis"
author: '[Guillermo Montero-Melis](https://www.mpi.nl/people/montero-melis-guillermo)'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
    code_folding: show
---

Introduction
============

This script analyzes the results as specified in our Registered Report,
see [here](https://osf.io/26fbh/).


Session set up
==============

Libraries:

```{r setup, warning = FALSE, message = FALSE}
library("knitr")
library("tidyverse")
library("brms")
library("egg")  # theme_article()
theme_set(theme_article())  # ggplot theme
knitr::opts_chunk$set(fig.height=4)
```

Source file where functions are defined:

```{r, message = FALSE}
source("myfunctions/analysis_fncs.R")
```

Load experiment data:

```{r}
d60_w_control <- read_csv("data60.csv") %>%
  rename(subject = ID_unique)
# Add nuisance variable "preced_error" (binary) which is 1 if an error was made
# on any of the preceding words in a trial, 0 otherwise:
preced_error <- function(x) {
  cumsum <- cumsum(x)
  shift_cumsum <- c(0, cumsum[1 : (length(cumsum) - 1)])
  out <- ifelse(shift_cumsum > 0, 1, 0)
  out
}
d60_w_control <- d60_w_control %>%
  group_by(subject, trial_exp) %>%
  mutate(preced_error = preced_error(error))
# data frame with only the critical conditions
d60 <- d60_w_control %>% filter(block_type != "CONTROL")
kable(head(d60))
```

Load the model fitted to the original data (see re-analysis - Appendix B)
and extract the posterior for the critical interaction effect for later use
as the prior for the replication Bayes factor:

```{r}
# Load brms model fitted to original data. It's a list with three objects: the
# full model is the first element of that list:
bfm_orig <- readRDS("sp13_bfm_max.rds")[[1]]
# Extract posterior estimates for the critical interaction effect from the brms
# model (pull converts it to a numerical vector):
beta_posterior <- pull(posterior_samples(bfm_orig, "b_mv_wt"))
sp13_posterior <- c(mean(beta_posterior), sd(beta_posterior))
names(sp13_posterior) <- c("M", "SD")
sp13_posterior
```


Visualization of raw data
========================

```{r}
d60_sbj <- d60_w_control %>%
  group_by(subject, block_type, word_type) %>%
  summarise(prop_error = mean(error))
```

Critical conditions:

```{r}
plot_raw_data(d60_w_control)
my_ggsave("SP13_replication_raw", 6, 4)
```

Including the control condition for comparison:

```{r}
plot_raw_data(d60_w_control, excl_CONTROL = FALSE)
my_ggsave("SP13_replication_raw_control", 6, 4)
```


Analysis with pre-registered analysis pipeline
========================================


Main analysis - with weakly informative priors
-------------------------------------------

Fit the models (or load if already fitted):

```{r}
ptm <- proc.time()
# analysis_pipe() tries to load the models from disk and fits them if that fails
results_brms_d60 <- analysis_pipe(d60)
# Stop the clock
time_models <- proc.time() - ptm
time_models
```

Model summaries:

```{r}
# summary(results_brms_d60[[1]])  # full
# summary(results_brms_d60[[2]])  # null
```


Compute Bayes Factor:

```{r}
ptm <- proc.time()
results_brms_d60_BF <- my_BF_wrapper(results_brms_d60)
time_bridge <- proc.time() - ptm
time_bridge
results_brms_d60_BF
1 / results_brms_d60_BF$bf
```


Secondary analysis - replication BF using original results from SP13 as priors
-------------------------------------------

Fit the models:

```{r}
ptm <- proc.time()
# results_brms_d60_repBF <- analysis_pipe(d60, repBF = TRUE, repBF_post = sp13_posterior)
# Stop the clock
time_models <- proc.time() - ptm
time_models
```


Model summaries:

```{r}
# summary(results_brms_d60_repBF[[1]])  # full
# summary(results_brms_d60_repBF[[2]])  # null
```


Compute Bayes Factor:

```{r}
# ...
```


Session info
============

```{r}
sessionInfo()
```

