---
title: "Analysis"
author: '[Guillermo Montero-Melis](https://www.mpi.nl/people/montero-melis-guillermo)'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
    code_folding: show
---

Introduction
============

This script analyzes the results at the pre-registered time points, i.e. after
collecting 60, 72, ... participants.


Session set up
==============

Libraries:

```{r setup, warning = FALSE, message = FALSE}
library("knitr")
library("tidyverse")
library("brms")
library("egg")  # theme_article()
theme_set(theme_article())  # ggplot theme
knitr::opts_chunk$set(fig.height=4)
```

Source file where functions are defined:

```{r, message = FALSE}
source("myfunctions/analysis_fncs.R")
```

Load data:

```{r}
d60_w_control <- read_csv("data60.csv") %>%
  rename(subject = ID_unique)
d60 <- d60_w_control %>% filter(block_type != "CONTROL")
```


Visualization of raw data
========================

```{r}
d60_sbj <- d60_w_control %>%
  group_by(subject, block_type, word_type) %>%
  summarise(prop_error = mean(error))
```


```{r}
my_dodge_w <- .2
my_jitter_w <- .2
my_alpha <- .4
d60_sbj %>%
  filter(block_type != "CONTROL") %>%
  mutate(block_type = as.factor(block_type)) %>%
  ggplot(
    aes(x = block_type, y = prop_error, color = word_type, shape = word_type)
    ) +
  geom_point(
    alpha = my_alpha,
    position = position_jitterdodge(jitter.width = my_jitter_w, jitter.height = 0, 
                                    dodge.width = my_dodge_w, seed = 6477)
    ) +
  stat_summary(
    fun.data = "mean_cl_boot", position = position_dodge(width = my_dodge_w),
    size = 1.5
  ) +
  stat_summary(
    aes(x = as.numeric(block_type)),
    fun = mean, 
    geom = "line",
    position = position_dodge(width = my_dodge_w),
    size = 1.25
    ) +
  xlab("Block type") +
  ylab("Proportion of errors") +
  labs(color = "word type", shape = "word type") +
  guides(color = guide_legend(override.aes = list(size = .75)))
my_ggsave("SP13_replication_raw", 6, 4)
```



```{r}
d60_sbj %>%
  filter(block_type != "CONTROL") %>%
  ggplot(aes(
    x = block_type, y = prop_error, color = word_type, shape = word_type)
    ) +
    geom_point(
      alpha = .3,
      position = position_jitterdodge(
        jitter.width = NULL, jitter.height = 0, dodge.width = 0.15, seed = 6477
        )
      ) +
    stat_summary(
      fun.data = "mean_cl_boot", position = position_dodge(width = .5)
      )
```

With the control condition for comparison



```{r}
my_dodge_w <- .2
my_jitter_w <- .2
my_alpha <- .4
d60_sbj %>%
  filter(block_type != "CONTROL") %>%
  mutate(block_type = as.factor(block_type)) %>%
  ggplot(
    aes(x = block_type, y = prop_error, color = word_type, shape = word_type)
    ) +
  geom_point(
    alpha = my_alpha,
    position = position_jitterdodge(jitter.width = my_jitter_w, jitter.height = 0, 
                                    dodge.width = my_dodge_w, seed = 6477)
    ) +
  stat_summary(
    fun.data = "mean_cl_boot", position = position_dodge(width = my_dodge_w),
    size = 1.5
  ) +
  # stat_summary(
  #   aes(x = as.numeric(block_type)),
  #   fun = mean, 
  #   geom = "line",
  #   position = position_dodge(width = my_dodge_w),
  #   size = 1.25
  #   ) +
  xlab("Block type") +
  ylab("Proportion of errors") +
  labs(color = "word type", shape = "word type") +
  guides(color = guide_legend(override.aes = list(size = .75)))
my_ggsave("SP13_replication_raw", 6, 4)
```



```{r}
d60_sbj %>%
  ggplot(aes(
    x = block_type, y = prop_error, color = word_type, shape = word_type)
    ) +
    geom_point(
      alpha = .3,
      position = position_jitterdodge(
        jitter.width = NULL, jitter.height = 0, dodge.width = 0.15, seed = 6477
        )
      ) +
    stat_summary(
      fun.data = "mean_cl_boot", position = position_dodge(width = .5)
      )
```


Minor processing
================

Add nuisance variable `preced_error` (binary) which is 1 if an error
was made on any of the preceding words in a trial, 0 otherwise:

```{r}
# Add nuisance variable "preced_error"
preced_error <- function(x) {
  cumsum <- cumsum(x)
  shift_cumsum <- c(0, cumsum[1 : (length(cumsum) - 1)])
  out <- ifelse(shift_cumsum > 0, 1, 0)
  out
}
d60 <- d60 %>%
  group_by(subject, trial_exp) %>%
  mutate(preced_error = preced_error(error))
```


Analysis with pre-registered analysis pipeline
========================================




Session info
============

```{r}
sessionInfo()
```

